### 2018-08-08 analysis/01_analysis.Rmd
shuffle trial sets

---
title: "Discon analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(langcog)
```

```{r data preprocessing}
# Data pre-processing

files <- dir("../../pilot_1_raw_data/production-results")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("../../pilot_1_raw_data/production-results/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$answers$data$data) %>%
    mutate(workerid = jd$WorkerId)
  raw_data <- bind_rows(raw_data, id)
}

data <- raw_data %>%
  mutate(correct_target3 = ifelse(correct_target1 == 0 & correct_target2 == 0, 1, 0)) %>%
  mutate(distribution = as.character(distribution),
         distribution = factor(distribution, levels = c("c(6, 0, 0)", "c(4, 2, 0)", "c(2, 2, 2)")),
         pick = ifelse(pickPos == "item_l", item_l, ifelse(pickPos == "item_r", item_r, ifelse(pickPos=="item_m", item_m, NA))),
         correct_item = ifelse(pick == correctItem, 1, 0),
         pickTest = substr(pick, 4, length(pick)),
         correct_target1 = ifelse(pickTest == target1, 1, 0),
         correct_target2 = ifelse(pickTest == target2, 1, 0),
         correct_target3 = ifelse(pickTest == target3, 1, 0))

write.csv(data, file="../data/pilot.data.csv")

# data <- read_csv(file="../data/pilot.data.csv")
```

```{r sanity checks}
data %>%
  filter(phase == 'training') %>%
  group_by(workerid) %>%
  summarise(correct = mean(correct_item))

drop <- data %>%
  filter(phase == 'training') %>%
  group_by(workerid) %>%
  summarise(correct = mean(correct_item)) %>%
  filter(correct < 0.83) %>%
  pull(workerid)

data <- data %>%
  filter(!workerid %in% drop)
```

```{r}
data %>%
  filter(phase == "test", distribution == "c(4, 2, 0)") %>%
  group_by(pickTest) %>%
  summarise(mean = mean(correct_target1))
```
# Do proportion of choices match the frequency in the distribution

```{r plot1}
plot1 <- data %>%
  filter(phase == 'test') %>%
  group_by(distribution, target1, workerid) %>%
  summarise(correct = mean(correct_target1),
            correct2 = mean(correct_target2),
            correct3 = mean(correct_target3)) %>%
  gather(item, prop_correct, -distribution, -workerid, -target1) %>%
  group_by(distribution, target1, item) %>%
  multi_boot_standard(col = "prop_correct")

ggplot(plot1,
       aes(x = item, y=mean, fill=item)) +
  geom_bar(stat='identity') +
#  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_grid(target1~distribution) +
  geom_hline(yintercept = 1/3, lty=2) +
  ylim(0, 1) +
  theme_few()

```

# Does the framing influence the match between choice and distribution

```{r plot 2}
plot2 <- data %>%
  filter(phase == 'test') %>%
  group_by(distribution, workerid, trialType) %>%
  summarise(correct = mean(correct_target1),
            correct2 = mean(correct_target2),
            correct3 = mean(correct_target3)) %>%
  gather(item, prop_correct, -distribution, -workerid, -trialType) %>%
  arrange(workerid)

 # group_by(distribution, trialType, item) %>%
 # multi_boot_standard(col = "prop_correct")

ggplot() + 
  geom_jitter(data = plot2, 
              aes(x = item, y=prop_correct, fill=item, col=trialType),
              height = 0.1,
              width = 0.1) +
  facet_grid(trialType~distribution) +
  theme(legend.position="none")

plot2 <- data %>%
  filter(phase == 'test') %>%
  group_by(distribution, workerid, trialType) %>%
  summarise(correct = mean(correct_target1),
            correct2 = mean(correct_target2),
            correct3 = mean(correct_target3)) %>%
  gather(item, prop_correct, -distribution, -workerid, -trialType) %>%
  group_by(distribution, trialType, item) %>% 
  multi_boot_standard(col = "prop_correct")

ggplot(plot2,
       aes(x = item, y=mean, fill=item)) +
  geom_bar(stat='identity') +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  facet_grid(trialType~distribution) +
  geom_hline(yintercept = 1/3, lty=2) +
  ylim(0, 1) +
  theme_few()

```

# Comparing the last input with the picked object by distribution

```{r plot 2}
a <- data %>%
  filter(slide == 6) %>%
  select(workerid, pick)

b <- data %>%
  filter(slide == "choice") %>%
  select(workerid, distribution, pickTest, correct_target1)

b$pickSlide6 = a$pick

b <- b %>%
  filter(distribution != "c(6, 0, 0)", correct_target1 == 1) %>%
  arrange(distribution) %>%
  group_by(distribution) %>%
  summarise(n = length(distribution))

# 30/59 trials picking the same category as the last input slide.
# 24/33 for 4, 2, 0
# 7/26 for 2, 2, 2

```

```{r}
data %>%
  filter(phase == "test", distribution == "c(6, 0, 0)") %>%
  group_by(pick) %>%
  summarise(n = length(correct_target1),
            mean = mean(correct_target1)) %>%
  arrange(mean)
```