### 2018-08-29 analysis/01_kids.Rmd
sound recordings

---
title: "DisCon Kids Pilot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(langcog)

```

```{r data pre-processing}

files <- dir("../../data/kids_pilot/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("../../data/kids_pilot/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$data$data)
  raw_data <- bind_rows(raw_data, id)
}

data <- raw_data %>%
  mutate(distribution = as.character(distribution),
         distribution = factor(distribution, levels = c("c(6, 0, 0)", "c(4, 2, 0)", "c(2, 2, 2)")))
```

```{r sanity checks}
data <- data %>% 
  filter(subid != "", subid != "test", subid != "Test")

data %>%
  filter(phase == 'training') %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_item))

drop <- data %>%
  filter(phase == 'training') %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_item)) %>%
  filter(correct < 0.83) %>%
  pull(subid)

data <- data %>%
  filter(!subid %in% drop)

```

# Do proportion of choices match the frequency in the distribution

```{r proportion of choices}
plot1 <- data %>%
  filter(phase == 'test') %>%
  group_by(distribution, subid) %>%
  summarise(correct = mean(correct_target1),
            correct2 = mean(correct_target2),
            correct3 = mean(correct_target3)) %>%
  gather(item, prop_correct, -distribution, -subid) %>%
  group_by(distribution, item) %>%
  multi_boot_standard(col = "prop_correct")

ggplot(plot1,
       aes(x = item, y=mean, fill=item)) +
  geom_bar(stat='identity') +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  scale_fill_solarized(name = "Chosen item",
                       breaks = c("correct", "correct2", "correct3"),
                       labels = c("most frequent", "second most frequent", "least frequent"))+
  facet_grid(~distribution, 
             labeller = as_labeller(c(`c(6, 0, 0)` = "Distribution\n(6-0-0)", `c(4, 2, 0)` = "Distribution\n(4-2-0)", `c(2, 2, 2)` = "Distribution\n(2-2-2)", preference = "preference", statistical = "statistical"))) +
  geom_hline(yintercept = 1/3, lty=2) +
  ylim(0, 1) +
  theme_few() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(x="", y= "proportion chosen")
```

# How does the last item seen affect the object chosen in the anaphora case? 
```{r last item}
plot2 <- data %>%
  filter(phase == "test", distribution != "c(6, 0, 0)") %>%
  mutate(trackLast = ifelse(lastInputCat == target1, "yes", "no")) %>%
  group_by (distribution, subid) %>%
  summarise (lastInput = mean(same_lastInput),
             n = n()) %>% 
  multi_boot_standard(col = "lastInput")

ggplot(plot2,
       aes(x = distribution, y = mean, fill = distribution)) +
  geom_bar(stat='identity', width = 0.3) +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_hline(yintercept = 1/3, lty=2) +
  ylim(0, 1) +
  theme_few()
```

# Are participants choosing the middle item more? 
```{r middle item}
plot3 <- data %>%
  filter(phase == 'test') %>%
  mutate(pick_l = ifelse(pickPos == "item_l", 1, 0), 
         pick_m = ifelse(pickPos == "item_m", 1, 0),
         pick_r = ifelse(pickPos == "item_r", 1, 0)) %>%
  group_by(distribution, subid) %>%
  summarise(choose_l = mean(pick_l),
            choose_m = mean(pick_m),
            choose_r = mean(pick_r)) %>%
  gather(item, prop_choose, -distribution, -subid) %>%
  group_by(distribution, item) %>%
  multi_boot_standard(col = "prop_choose")

ggplot(plot3,
       aes(x = item, y=mean, fill=item)) +
  geom_bar(stat='identity') +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  scale_fill_solarized(name = "Chosen item",
                       breaks = c("choose_l", "choose_m", "choose_r"),
                       labels = c("left", "middle", "right"))+
  facet_grid(~distribution, 
             labeller = as_labeller(c(`c(6, 0, 0)` = "Distribution\n(6-0-0)", `c(4, 2, 0)` = "Distribution\n(4-2-0)", `c(2, 2, 2)` = "Distribution\n(2-2-2)"))) +
  geom_hline(yintercept = 1/3, lty=2) +
  ylim(0, 1) +
  theme_few() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(x="", y= "proportion chosen")
```

# Are participants choosing the middle item even when the middle item is wrong? 
```{r middle item, chose incorrectly}
plot4 <- data %>%
  filter(phase == 'test', correct_target1 == 0) %>%
  mutate(pick_l = ifelse(pickPos == "item_l", 1, 0), 
         pick_m = ifelse(pickPos == "item_m", 1, 0),
         pick_r = ifelse(pickPos == "item_r", 1, 0)) %>%
  group_by(distribution, subid) %>%
  summarise(choose_l = mean(pick_l),
            choose_m = mean(pick_m),
            choose_r = mean(pick_r)) %>%
  gather(item, prop_choose, -distribution, -subid) %>%
  group_by(distribution, item) %>%
  multi_boot_standard(col = "prop_choose")

ggplot(plot4,
       aes(x = item, y=mean, fill=item)) +
  geom_bar(stat='identity') +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  scale_fill_solarized(name = "Chosen item",
                       breaks = c("choose_l", "choose_m", "choose_r"),
                       labels = c("left", "middle", "right"))+
  facet_grid(~distribution, 
             labeller = as_labeller(c(`c(6, 0, 0)` = "Distribution\n(6-0-0)", `c(4, 2, 0)` = "Distribution\n(4-2-0)", `c(2, 2, 2)` = "Distribution\n(2-2-2)"))) +
  geom_hline(yintercept = 1/3, lty=2) +
  ylim(0, 1) +
  theme_few() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(x="", y= "proportion chosen")
```