### 2018-10-30 analysis/02_kids_power_analysis_600.Rmd
add power analysis for kids 600 4 cat

---
title: "DisCon Kids Pilot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(langcog)

library(lsr)
library(pwr)

```

```{r data}
# data from pilot
data <- read_csv(file="../data/kids_4cat_pilot.data.csv")
```
# Data processing
```{r sanity checks}
data <- data %>% 
  filter(subid != "", subid != "test", subid != "Test")

# check training performance
data %>%
  filter(phase == 'training') %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_item))

# filter out children that don't get known words right
drop <- data %>%
  filter(phase == 'training') %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_item)) %>%
  filter(correct < 0.83) %>%
  pull(subid)

data %>%
  filter(!subid %in% drop) %>%
  group_by(subage) %>%
  summarise(n = length(unique(subid)))

```

# Visualizing the data

```{r}
p1 <- data %>%
      filter(phase == "test")%>% # only test trials
  group_by(subage, subid) %>%
  summarise(correct = mean(correct_target1))

p2 <- p1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p1, aes(x = subage, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = subage, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="Proportion Correct Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Age")
```

# Power analysis
``` {r power analysis}
power <- data %>%
  filter(phase == 'test') %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_target1))

t.test(power$correct, mu = 1/3)

d <- cohensD(power$correct, mu = 1/3)

pwr.t.test(d = d, sig.level = 0.05, power = .8, type = c("one.sample"))
```