### 2019-01-18 analysis/01_adult_speakerchange.Rmd
data file for speaker change preference experiment

---
title: "DisCon Speaker Change Adult Analysis"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(langcog)
```

#Pilot 1 data
```{r data preprocessing}
# Data pre-processing

files <- dir("../../adult_600_speakerchange/production-results/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("../../adult_600_speakerchange/production-results/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  id <- as_data_frame(jd$answers$data$data) %>%
    mutate(workerid = jd$WorkerId)
  raw_data <- bind_rows(raw_data, id)
}

data <- raw_data %>%
  mutate(distribution = as.character(distribution),
         distribution = ifelse(distribution == "2:0", "c(2, 1, 0)", distribution),
         distribution = factor(distribution, levels = c("c(6, 0, 0)", "c(4, 2, 0)", "c(2, 2, 2)", "c(3, 0, 0)", "c(1, 1, 1)", "c(2, 1, 0)")))

```

```{r}
data_ip <- data %>%
  filter(phase == 'training') %>%
  group_by(workerIp) %>%
  summarise(correct = mean(correct_item))
```

```{r sanity checks}
data %>%
  filter(phase == 'training') %>%
  group_by(workerid) %>%
  summarise(correct = mean(correct_item))

drop <- data %>%
  filter(phase == 'training') %>%
  group_by(workerid) %>%
  summarise(correct = mean(correct_item)) %>%
  filter(correct < 0.83) %>%
  pull(workerid)

data <- data %>%
  filter(!workerid %in% drop)
```
```{r}

sample_size <- data %>%
  distinct(workerid, .keep_all = TRUE)

sample_size %>%
 group_by(experiment) %>%
 summarise(n = n())
```

```{r plot1}
plot1 <- data %>%
  filter(phase == 'test') %>%
  group_by(distribution, workerid, speakerChange) %>%
  summarise(correct = mean(correct_target1),
            correct2 = mean(correct_target2),
            correct3 = mean(correct_target3)) %>%
  gather(item, prop_correct, -distribution, -workerid, -speakerChange) %>%
  group_by(distribution, item, speakerChange) %>%
  multi_boot_standard(col = "prop_correct")

ggplot(plot1,
       aes(x = item, y=mean, fill=item)) +
  geom_bar(stat='identity') +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  scale_fill_solarized(name = "Chosen item",
                       breaks = c("correct", "correct2", "correct3"),
                       labels = c("most frequent", "second most frequent", "least frequent"))+
  facet_grid(~speakerChange, 
             labeller = as_labeller(c(change = "speaker change", same = "no change"))) +
  geom_hline(yintercept = 1/3, lty=2) +
  ylim(0, 1) +
  theme_few() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(x="", y= "proportion chosen")

plot3 <- data %>%
  filter(phase == 'test') %>%
  group_by(distribution, workerid, speakerChange) %>%
  summarise(correct = mean(correct_target1)) %>%
  gather(item, prop_correct, -distribution, -workerid, -speakerChange) %>%
  group_by(distribution, item, speakerChange)

plot3a <- plot3 %>%
  multi_boot_standard(col = "prop_correct")

ggplot() +
  geom_jitter(data = plot3, aes(x = speakerChange, y = prop_correct, col = speakerChange, alpha = .2), width = .3,height = .025)+
  geom_pointrange(data = plot3a, aes(x = speakerChange, y = mean, col = speakerChange, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="Speaker Change", y="Correct Item Chosen")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F) +
scale_colour_solarized(name = "Speaker Change",
                       breaks = c("correct", "correct2", "correct3"),
                       labels = c("most frequent", "second most frequent", "least frequent"))

```
```{r wilcoxon test}
## Wilcoxon tests against chance
library(exactRankTests)

data %>%
  filter(phase == "test") %>%
  group_by(distribution, workerid) %>%
  summarise(correct = mean(correct_target1)) %>%
  summarise(correct = list(correct)) %>%
  group_by(distribution) %>%
  mutate(mean= mean(unlist(correct)),
         stat = wilcox.exact(unlist(correct), mu = 0.5)$statistic,
         p = wilcox.exact(unlist(correct), mu = 0.5)$p.value) %>%
  select(distribution,mean,stat,p) %>%
  knitr::kable(digits = 3)
```

```{r Bayesian}
library(brms)

data_inf <- data%>%
  filter(phase == "test")

prior1 <- get_prior(correct_target1 ~ speakerChange * experiment +
                      (speakerChange | workerid) + (speakerChange | agent),
                    data = data_inf, family = bernoulli())

# registered  model
bm1 <- brm(correct_target1 ~ speakerChange * experiment +
                      (speakerChange | workerid) + (speakerChange | agent),
                    data = data_inf, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          save_all_pars = TRUE,
          prior = prior1,
          iter = 20000)
summary(bm1)

prior2 <- get_prior(correct_target1 ~ speakerChange +
                      (speakerChange | workerid) + (speakerChange | agent),
                    data = data_inf, family = bernoulli())

bm2 <- brm(correct_target1 ~ speakerChange  +
                      (speakerChange | workerid) + (speakerChange | agent),
                    data = data_inf, family = bernoulli(),
          control = list(adapt_delta = 0.99),
          sample_prior = F,
          save_all_pars = TRUE,
          prior = prior2,
          iter = 20000)
summary(bm2)

saveRDS(bm1, "01_adult_speakerchange_preference_models/bm1.rds")
saveRDS(bm2, "01_adult_speakerchange_preference_models/bm2.rds")

bf_full_null <- bayes_factor(bm1, bm2, log = FALSE, maxiter = 5000)

bf_full_null
```