### 2019-03-01 analysis/02_kids_600.Rmd
add kids data

---
title: "DisCon Kids Pilot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(jsonlite)
library(tidyr)
library(stringr)
library(dplyr)
library(ggthemes)
library(langcog)
library(readxl)
library(knitr)

library(lsr)
library(pwr)

library(coda)
library(reshape2)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

```

```{r data pre-processing}

#files <- dir("../../kids_simple_4cat/")
files <- dir("../../kids_raw_data/600/")

raw_data <- data_frame()
for (f in files) {
#  jf <- paste("../../kids_simple_4cat/",f,sep="")
  jf <- paste("../../kids_raw_data/600/",f,sep="")
  jd <- fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,36,str_length(jf)-11)
  id <- as_data_frame(jd$data$data)%>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

#unique(raw_data$subid)

# linking the data to the demographics spreadsheet
# always check if subjects get "lost" due do differetn subids
log <- read_excel("../../DisCon-subject_log.xlsx", 1)%>%
  filter(Condition == "d600")%>%
  mutate(dob = as.numeric(dob),
         dob = as.Date(dob,origin = "1899-12-30"))%>%
  select(subid,experimenter,keep_drop,sex,dob)

#unique(log$subid)

data <- left_join(raw_data,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(distribution = as.character(distribution),
         distribution = factor(distribution, levels = c("c(6, 0, 0)", "c(4, 2, 0)", "c(2, 2, 2)")),
         age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))

#unique(data$subid)

#write.csv(data, file="../data/kids_600.data.csv")

data <- read_csv(file="../data/kids_600.data.csv")
```

```{r sanity checks}
# filtering out subjects that perform poorly on training trials
data %>%
  filter(phase == 'training') %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_item))

drop <- data %>%
  filter(phase == 'training') %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_item)) %>%
  filter(correct < 0.83) %>%
  pull(subid)

data <- data %>%
  filter(!subid %in% drop)
```

# Participants

```{r}
data %>%
  group_by(subage)%>%
  summarise(n = length(unique(subid)))%>%
  knitr::kable(digits = 2)
```

# Proportion correct choice in training trials

```{r}
pt1 <- data %>%
  filter(phase == 'training')%>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct_item))

pt2 <- pt1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = pt1, aes(x = subage, y = correct, col = subage, alpha = .2),width = .1,height = .025)+
  geom_pointrange(data = pt2, aes(x = subage, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="Proportion correct in training")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  scale_color_solarized()
```

# Proportion correct choice in test trials across age bins

```{r proportion of choices}
p1 <- data %>%
  filter(phase == 'test')%>%
  group_by(experiment,subid) %>%
  summarise(correct = mean(correct_target1))

p2 <- p1 %>%
  group_by(experiment) %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p1, aes(x = experiment, y = correct,  alpha = .2),width = .2,height = .01)+
  geom_pointrange(data = p2, aes(x = experiment, y = mean,  ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="Proportion correct in training")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  scale_color_solarized()
```

# Proportion correct choice in test trials by age bin

```{r proportion of choices}
p3 <- data %>%
  filter(phase == 'test')%>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct_target1))

p4 <- p3 %>%
  group_by(subage) %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p3, aes(x = subage, y = correct, col = subage, alpha = .2),width = .2,height = .01)+
  geom_pointrange(data = p4, aes(x = subage, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  scale_color_solarized(name = "Age")
```

Proportion correct choice by age continous

```{r}
p5 <- data %>%
  filter(phase == 'test')%>%
  group_by(age_num, subid) %>%
  summarise(correct = mean(correct_target1))

ggplot(data = p5, aes(x = age_num, y = correct)) +
  geom_jitter(width = .01, height = .0, alpha = .5)+
  geom_smooth(method = "lm",  se = T,  alpha = .5, size = 1, col = "darkred", fill = "darkgreen", alpha = .5 )+
  geom_hline(yintercept = 1/3, lty=2)+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(3,5)+
  guides(alpha = F)
```

# Bayesian t-test against chance across age groups

```{r}
## one sample bayesian t-tests
library(BayesFactor)

data %>%
  filter(phase != "training") %>%
  group_by(subid) %>%
  summarise(correct = mean(correct_target1)) %>%
  summarise(correct = list(correct)) %>%
  mutate(mean= mean(unlist(correct)),
         bf = extractBF(ttestBF(unlist(correct), mu = 1/3))$bf) %>%
  select(mean,bf) %>%
  knitr::kable(digits = 3)

```

# BAyesian t-test by age group

```{r}
## one sample bayesian t-tests
library(BayesFactor)

data %>%
  filter(phase != "training") %>%
  group_by(subage,subid) %>%
  summarise(correct = mean(correct_target1)) %>%
  summarise(correct = list(correct)) %>%
  group_by(subage) %>%
  mutate(mean= mean(unlist(correct)),
         bf = extractBF(ttestBF(unlist(correct), mu = 1/3))$bf) %>%
  select(subage,mean,bf) %>%
  knitr::kable(digits = 3)

```
Plotting posterior based on t-test

```{r}
plot_post <- data %>%
  filter(phase != "training") %>%
  group_by(subage,subid) %>%
  summarise(correct = mean(correct_target1))%>%
  summarise(correct = list(correct)) %>%
  group_by(subage) %>%
  mutate(posterior = list((ttestBF(unlist(correct), mu = 1/3, posterior = T, iterations = 1000)[,"mu"])))%>%
  select(-correct)

plot_dens_ci <- plot_post %>%
  summarise(mode = estimate_mode(unlist(posterior)),
            lci = hdi_lower(unlist(posterior)),
            uci = hdi_upper(unlist(posterior)))

#### make density plot ?

plot_dens <- data_frame(
  age = rep(c(3,4), each = 1000),
  posterior = c(unlist(plot_post%>%filter(subage == "3") %>% pull(posterior)),unlist(plot_post%>%filter(subage == "4") %>% pull(posterior)))
)

densities <- plot_dens %>%
  group_by(age) %>%
  do(., dens = density(.$posterior))

densities_df <-
  data.frame(
    age = c(rep('3', 512),
              rep('4', 512)),
    x = c(densities$dens[[1]]$x, 
          densities$dens[[2]]$x),
    y = c(densities$dens[[1]]$y,
          densities$dens[[2]]$y)
  ) %>%
  group_by(age)%>%
  mutate (
    ci_shade = ifelse( age == "3",
      ifelse( x <= plot_dens_ci%>%filter(subage == "3")%>%pull(lci) | x >= plot_dens_ci%>%filter(subage == "3")%>%pull(uci), "OFF", "ON"), ifelse( x <= plot_dens_ci%>%filter(subage == "4")%>%pull(lci) | x >= plot_dens_ci%>%filter(subage == "4")%>%pull(uci), "OFF", "ON")))

ggplot(plot_dens, aes(posterior, group = factor(age))) +
  geom_density(alpha = .5)+
  geom_vline(xintercept = 1/3, lty=2)+
  geom_area(data = densities_df %>% filter(ci_shade == "ON"), aes(x = x, y = y, fill = age), alpha = .5) + 
  labs(x="Proportion Correct")+
  theme_few() +
  xlim(-0.05,1.05)+
  guides(alpha = F)+
  scale_fill_solarized(name = "Age")+
  theme(axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

```

# Model 
```{r}
library(brms)

model_data <- data%>%
  filter(phase == "test")%>%
  mutate(age_num = scale(age_num))

#   model

kbm1 <- brm(correct_target1 ~ age_num +
                      (1| subid) + (1 | target1),
                    data = model_data, family = bernoulli(),
          control = list(adapt_delta = 0.95),
          sample_prior = F,
          save_all_pars = TRUE,
          iter = 2000)

summary(kbm1)
```

```{r}
bayesplot::mcmc_areas(as.array(kbm1), pars = c("b_Intercept", "b_age_num"), prob = 0.95)+
  geom_vline(xintercept = 0, lty = 2, col = "black")+
  theme_few()
```

```{r}
kbm2 <- brm(correct_target1 ~ 1 +
                      (1| subid) + (1 | target1),
                    data = model_data, family = bernoulli(),
          control = list(adapt_delta = 0.95),
          sample_prior = F,
          save_all_pars = TRUE,
          iter = 2000)

summary(kbm2)
```

## Model comparison

```{r}
child_waic <- brms::WAIC(kbm1, kbm2, compare = F)

child_weights <- model_weights(kbm1, kbm2)

# Bayes Factors (slightly variable)
bf_full_null <- bayes_factor(kbm1, kbm2, log = FALSE, maxiter = 5000)
```

```{r}
data_frame(
  model = c("kbm1_w_age","kbm2"),
  WAIC = c(child_waic$kbm1$estimates[3,1],child_waic$kbm2$estimates[3,1]),
  SE = c(child_waic$kbm1$estimates[3,2],child_waic$kbm2$estimates[3,2]),
  weight = c(child_weights[1],child_weights[2]),
  BF_bm1_bm2 = c(round(bf_full_null$bf,2),"-"))%>%
  kable(digits = 2)
```